Simulations of Ballot Polling Risk-Limiting Audits

### 2022-10-10 providence/sections/intro.tex
rewrote intro
kept a copy of the original intro

%Intro
The literature contains numerous descriptions of vulnerabilities in deployed voting systems, and it is not possible to be certain that any system, however well-designed, will perform as expected in all instances. For this reason, 
{\em evidence-based elections} \cite{evidence-based} aim to produce trustworthy and compelling evidence of the correctness of election outcomes, enabling the detection of problems with high probability. One way to implement an evidence-based election is to use a well-curated voter-verified paper trail, compliance audits, and a rigorous tabulation audit of the election outcome, known as a risk-limiting audit (RLA) \cite{RLA}. An RLA is an audit which guarantees that the probability of concluding that an election outcome is correct, given that it is not, is below a pre-determined value known as the risk limit of the audit, independent of the true, unknown vote distribution of the underlying election. Over a dozen states in the US have seriously explored the use of RLAs---some have pilot programs, some allow RLAs to satisfy a general audit requirement and some have RLAs in statute.

This paper concerns ballot-polling RLAs, which require a large number of ballots relative to comparison RLAs but do not rely on any special features of the election technology. Since comparison RLAs are not always feasible, ballot-polling audits remain an important resource and have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania and elsewhere). In the general ballot-polling RLA, a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}. A statistical measure is then computed to determine whether there is sufficient evidence to declare the election outcome correct within the pre-determined risk limit. Because the decision is made after drawing a round of ballots, the audit is termed a {\em round-by-round (R2)} audit. The special case when round size is one---that is, stopping decisions are made after each ballot draw---is a {\em ballot-by-ballot (B2)} audit.

The \BRAVO audit is designed for use as a B2 audit: it requires the smallest expected number of ballots when the true tally of the underlying election is as announced and stopping decisions are made after each ballot draw. In practice, election officials draw many ballots at once, and the \BRAVO stopping rule needs to be modified for use in an R2 audit that is not B2. There are two obvious approaches. The B2 stopping condition can be applied once at the end of each round: End-of Round (EoR) \BRAVO.  Alternatively, the order of ballots in the sample can be tracked by election officials and the B2 \BRAVO stopping condition can be applied retroactively after each ballot drawn: Selection-Ordered (SO) \BRAVO. SO \BRAVO requires fewer ballots on average than EoR \BRAVO but requires the work of tracking the order of ballots rather than just their tally.

\Minerva was designed for R2 audits and applies its stopping rule once for each round. Thus it does not require the tracking of ballots that SO \BRAVO does. Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced.
%\footnote{Their proof assumes that the number of relevant ballots drawn in each round is know beforehand. In MINERVA, the number of ballots drawn in each round is determined before any ballots are drawn. Because invalid ballots and ballots that are inconsequential for the contest being audited would be drawn in addition to relevant ballots, the assumption used by the proof is not true in general. (We are grateful to Philip Stark for drawing our attention to this.) However, any variation in number of relevant ballots drawn for a fixed round size would be random and not chosen by an adversary; the proof showing the risk-limiting property of MINERVA could hence be extended.}
They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for first round sizes with a large probability of stopping when the (true) underlying election is as announced. 
Broadrick {\em et al.} provide further simulations that show \Minerva requires fewer ballots over multiple rounds and for lower stopping probability.

While more efficient than \BRAVO, \Minerva requires that the round schedule is fixed in advance of the audit. This may lead to significant unnecessary work. For example, suppose the fixed round schedule for \Minerva is $10,000$ ballots per round (a reasonable number in a state-wide contest with a narrow margin), and the first round sample finds a number of ballots for the winner just short of that necessary for the audit to stop. It then may be sufficient to draw a very small second round and still stop with high probability in the second round, but the additional $10,000$ ballots must be drawn. In contrast, subsequent \BRAVO round sizes can be chosen based on preceding samples.

An open question is whether a ballot polling RLA exists with the efficiency of \Minerva and this flexibility of \BRAVO.

\subsection{Our Contributions}
We present \Providence, and provide the following:
\begin{enumerate}
\item Proof that \Providence is an RLA and resistant to an adversary who can choose subsequent round sizes with knowledge of previous samples
\item Simulations of \Providence, \Minerva, SO \BRAVO, and EoR \BRAVO which show that \Providence has similar efficiency to \Minerva, both greater than either implementation of \BRAVO
\item Results and analysis from the use of \Providence in a pilot audit in Rhode Island
\item Open source implementation of \Providence 
%including the novel metric Probability of Misleading(name?)
\end{enumerate}

>>>>

%Intro
It is well-known that electronic voting systems are vulnerable to software errors and manipulation which may be undetected. Errors and/or manipulation may not always change an election outcome, but we would want to know when they do. {\em Software independent} voting systems \cite{SI-Wack,rivest2008notion} are ones where an undetected change in the software cannot lead to an undetected change in the election outcome. {\em Evidence-based elections} \cite{evidence-based} use software independent systems to produce trustworthy evidence of outcome correctness; incorrect outcomes are detected with high probability when the evidence is examined. One approach to evidence-based elections is to use voter-verified paper ballots, store them securely, and perform public audits---a compliance audit to determine whether the ballots were stored securely; and a rigorous tabulation audit, known as a risk-limiting audit (RLA) \cite{RLA}, to determine whether the outcome is correctly computed from the stored ballots. Many US states have had pilot RLA programs. Additionally, some states allow RLAs to be used towards audit requirements, and some states require RLAs before elections can be certified.

We propose the \Providence audit, a new approach to the ballot polling RLA, and propose a new model for the work load of an election. We show that \Providence is superior to the popular ballot polling RLA \Bravo for real elections, and describe the use of our open source implementation by the Rhode Island Board of Elections for an audit of their 2021 elections. Our implementation of \Providence is likely to be useful later this year; ballot polling audits are expected to be used as pre-certification RLAs in at least one statewide contest in both Georgia and Pennsylvania for the 2022 general elections in the US.

\subsection{Background}
Ballot comparison RLAs require the fewest ballots of all known RLA approaches. On the other hand, they require the use of special election technology and are not always feasible. Ballot-polling RLAs require a larger number of ballots, but are more feasible because they do not require any additional functionality of the voting system. What is needed is a complete ballot manifest (a list of ballot storage containers and the number of ballots in each) which enables the creation of a well defined list of the ballots and their locations (the fifth ballot in box number 20, for example).

A ballot polling audit begins when a {\em round} \cite{usenix_minerva} of multiple randomly-chosen ballots is drawn. A risk measure is then computed to determine whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn. Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds. Ballot polling audits have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere) and in real statewide audits (Georgia, Virgina) \cite{vv_audits} as well as in audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}.

A {\em round-by-round (R2)} audit is one where the decision of whether to draw more ballots or not is taken after drawing a round of ballots; typically hundreds or thousands or tens of thousands of ballots in statewide elections. A {\em ballot-by-ballot (B2)} audit is the special case of round size one---when the decision is made after each ballot is drawn. The popular \BRAVO audit requires the smallest expected number of ballots when the announced tally of the election is correct, and stopping decisions are taken a ballot at a time (that is, when it is used as a B2 audit). Election officials typically draw ballots in large round sizes, and \BRAVO needs to be modified for use in this manner. For use as an R2 audit, the \BRAVO stopping condition can be applied once at the end of each round (End-of Round (EoR)), or retroactively after each ballot drawn if ballot order is retained (Selection-Ordered (SO)). SO \BRAVO is closer to the original B2 \BRAVO, and requires fewer ballots on average than EoR \BRAVO. But it requires the additional effort of tracking the order of ballots, and not simply the sample tally.

Zag\'{o}rski {\em et al.} propose ballot polling RLA \Minerva \cite{usenix_minerva}, which uses only the sample and round tallies, and does not need ballot order. They prove that it requires fewer ballots than EoR \BRAVO when both audits have the same pre-determined round schedule and the true tally is as announced. 
%\footnote{Their proof assumes that the number of relevant ballots drawn in each round is know beforehand. In MINERVA, the number of ballots drawn in each round is determined before any ballots are drawn. Because invalid ballots and ballots that are inconsequential for the contest being audited would be drawn in addition to relevant ballots, the assumption used by the proof is not true in general. (We are grateful to Philip Stark for drawing our attention to this.) However, any variation in number of relevant ballots drawn for a fixed round size would be random and not chosen by an adversary; the proof showing the risk-limiting property of MINERVA could hence be extended.}
They also present first-round simulations demonstrating that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first round sizes when the true tally is as announced. 
Broadrick {\em et al.} provide further simulations that show \Minerva requires fewer ballots over multiple rounds and for lower stopping probability \cite{simulations}, though the improvement from using \Minerva over either version of \BRAVO decreases with round size.

A major limitation of \Minerva is that one needs to determine the round schedule before the audit begins, because \Minerva has not been shown to be risk-limiting if an adversary can choose subsequent round sizes after knowing the sample drawn. \BRAVO, on the other hand, is not limited in this manner. This allows \BRAVO audits the flexibility of choosing smaller subsequent round sizes if the sample drawn so far is a ``good'' sample. An open question is whether a ballot polling RLA exists with the efficiency of \Minerva and this flexibility of \BRAVO.

A major limitation of our understanding of the ballot polling problem as a community is that we use the number of ballots drawn or values proportional to this number \cite{mclaughlin_thesis,bernhard-diss,RI-report} as measures of the workload of an audit. If this were a correct measure of the workload of an audit, we would want to use B2 audits (round size is one) and make decisions about stopping the audit after drawing each ballot, because this leads to the smallest expected number of ballots. Election officials, on the other hand, greatly prefer drawing many ballots at once. This preference is likely because each round has an overhead workload as well, including setting up the round and communicating among the various localities involved in conducting the audit (for example, audits of statewide contests involve the drawing of ballots at county offices where the ballots are stored). Further, there is an overhead to finding a storage box and opening it. Finally, in the current environment of misinformation, election officials would want to ensure that the probability of a misleading audit sample (falsely indicating that the loser won) is very small, which implies that round sizes should be large. Thus the workload of an audit is not simply a linear (or affine) function of the number of ballots drawn. Relatedly, an optimal round schedule is not completely determined by the expected number of ballots drawn. It depends on other variables as well.

\subsection{Our Contributions}
We present \Providence, and provide the following:
\begin{enumerate}
\item Proof that \Providence is an RLA and resistant to an adversary who can choose subsequent round sizes with knowledge of previous samples.
\item Simulations of \Providence, \Minerva, SO \BRAVO, and EoR \BRAVO which show that \Providence uses number of ballots similar to those of \Minerva, both fewer than either version of \BRAVO.
\item Results and analysis from the use of \Providence in a pilot audit in Rhode Island.
\item Open source implementation of \Providence. 
\item A model of workload that includes the overhead effort of each round and the overhead effort of retrieving a storage unit of ballots; simulations that illustrate the use of this model to compare the different types of ballot polling audits and to plan an audit with minimal workload.
\item An analysis of round size as a function of the maximum acceptable probability of a misleading audit sample.
%including the novel metric Probability of Misleading(name?)
\end{enumerate}

Our results demonstrate the superiority of \Providence over the other audits. We hope that our work will help election officials better plan ballot polling audits, including in Georgia and Pennsylvania in 2022.
